Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-09 papers/FX/draft.tex
some progress on the student's t case.
close to finishing the proof in the Student's t case.

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
Then it follows
\begin{eqnarray*}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
}
\end{eqnarray*}
The nummerator of the right side of $\pd{f(x, \nu)}{\nu}$, if seen as
a function of $x^2$, has a single root:
\begin{equation}
  \label{eq:xxie5}
  x^2 = \nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu
\end{equation}
where $W(\cdot)$ is the principle branch of the Lambert $W$
function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
the right side of \eqref{eq:xxie5} for positivity, we first note
$c'(\nu) > 0$:
\[
c'(\nu) = {
  \pi \Gamma(\nu/2 + 1/2) \left[
    \nu \Psi(\nu/2 + 1/2) - \nu \Psi(\nu/2) - 1
  \right]
  \over
  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
}
\]
where $\Psi(\cdot)$ is the digamma function:
\[
\Psi(x) = \td{\ln[\Gamma(x)]}{x}
\]
Since for all $x > 0$, $\Psi(x) > 0$ and $\nu > 1$, it is clear
$c'(\nu) > 0$. Furthermore, we recall
$W(\cdot)$ is increasing on its principle branch. So
\begin{eqnarray*}
  &&
  W\left[
    -\left( 1 + {1 \over \nu} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &>& 
  W\left[
    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &=& W(-y e^{-y}) + y
\end{eqnarray*}
where
\[
y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
\]
Now notice
\[
\ln(y e^{-y}) = \ln(y) - y
\]
is a decreasing function. Thus $-y e^{-y}$ is an increasing
function. Hence we have
\begin{eqnarray*}
  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
\end{eqnarray*}
Now it is clear
\[
\nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu > 0
\]

>>>>

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
Then it follows
\begin{eqnarray}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  } \nonumber \\
  &:=& {
    P(x^2, \nu)
    \over
    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  }
  \label{eq:xxie4.1}
\end{eqnarray}

While the denominator of the right side of $\pd{f(x, \nu)}{\nu}$ is
always positive, its nummerator $P(x^2, \nu)$ has a single root:
\begin{equation}
  \label{eq:xxie5}
  x_0^2 = \nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu
\end{equation}
where $W(\cdot)$ is the principle branch of the Lambert $W$
function. and $c'(\cdot)$ is the derivative of $c(\cdot)$. To check
the right side of \eqref{eq:xxie5} for positivity, we first note
$c'(\nu) > 0$:
\[
c'(\nu) = {
  \pi \Gamma(\nu/2 + 1/2) \left\{
    \nu \left[ \Psi(\nu/2 + 1/2) - \Psi(\nu/2)\right] - 1
  \right\}
  \over
  2 \Gamma(\nu/2) (\pi \nu)^{3/2}
}
\]
where $\Psi(\cdot)$ is the digamma function:
\[
\Psi(x) = \td{\ln[\Gamma(x)]}{x}
\]
\begin{minipage}{0.48\textwidth}
  $\Psi(x)$ \linebreak
  \includegraphics[width=\textwidth]{digamma.png}
\end{minipage}\hfill
\begin{minipage}{0.5\textwidth}
  As shown in the figure to the left, $\Psi(x)$ is increasing
  for $x > 0$. Therefore $\Psi(\nu/2 + 1/2) - \Psi(\nu/2) > 0$.
  So we have
  \begin{eqnarray*}
    && \nu \left[
    \Psi(\nu/2 + 1/2) - \Psi(\nu/2)
    \right] - 1 \\
    &\geq& 1 \times \left[
    \Psi(1/2 + 1/2) - \Psi(1/2)
    \right] - 1 \\
    &=& \ln(4) - \ln(e) \\
    &>& 0
  \end{eqnarray*}
  Thus $c'(\nu) > 0$. Furthermore, we recall
  $W(\cdot)$ is increasing on its principle branch. So
\end{minipage}
\begin{eqnarray*}
  &&
  W\left[
    -\left( 1 + {1 \over \nu} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &>& 
  W\left[
    -\left( 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \right)
    e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
  \right]
  + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} \\
  &=& W(-y e^{-y}) + y
\end{eqnarray*}
where
\[
y = 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)} > 1
\]
Now notice
\[
\ln(y e^{-y}) = \ln(y) - y
\]
is a decreasing function. Thus $-y e^{-y}$ is an increasing
function. Hence we have
\begin{eqnarray*}
  W(-y e^{-y}) + y &>& W(-e^{-1}) + 1 = 0
\end{eqnarray*}
Now it is clear
\[
\nu\exp\left\{
    W\left[
      -\left(1 + {1 \over \nu}\right)
      e^{-1 - 2 c'(\nu)/c(\nu) - 1/\nu}
    \right]
    + 1 + {1 \over \nu} + {2 c'(\nu) \over c(\nu)}
  \right\} - \nu > 0
\]
Now that we have established that $\pd{f}{\nu}(x, \nu) = 0$ has a
single positive root, it remains to determine the sign of
$\pd{f}{\nu}(x, \nu)$ on the two sides of the root in order to apply
lemma \ref{lemma:I}. For this purpose we observe
\begin{equation}
  \label{eq:xxie5.1}
  P(0, \nu) = 2 \nu^2 c'(\nu) > 0
\end{equation}
So we want to investigate $\pd{P}{x^2}(x^2, \nu)$:
\begin{eqnarray}
  && \pd{P}{x^2}(x^2, \nu) \nonumber \\
  &=&
  2 \nu c'(\nu) + c(\nu) - \nu c(\nu) \ln\left(
  1 + {x^2 \over \nu}
  \right)
  \label{eq:xxie6}
\end{eqnarray}
From \eqref{eq:xxie5.1} and \eqref{eq:xxie6} it is clear
$P(x^2, \nu) > 0$ for $0 \leq x^2 < x_0^2$ and
$P(x^2, \nu) < 0$ for $x^2 > x_0^2$.
